---
title: "curatedENCORE: interfaces to ENCORE for RBP biology in AnVIL"
shorttitle: "outlier dytection methods"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{curatedENCORE: interfaces to ENCORE for RBP biology in AnVIL}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction

This package provides examples of exploring and using resources in
the ENCORE project resources made available in NHGRI AnVIL.

We address the following concerns:

- What ENCORE resources are available in AnVIL, and how can
the Bioconductor AnVIL package be used to survey and
retrieve them?

- How can ENCORE resources be combined with other genomic
reference data and experiments to foster discovery in RNA biology
and its application to disease mechanisms?

# Basic reference information

We inferred the names of genes coding for RNA-binding proteins (RBPs)
from filename patterns in ENCORE.  The genomic coordinates for these
genes are available in `rbp_cands_gr`.

```{r lkbasic, message=FALSE}
library(curatedENCORE)
data(rbp_cands_gr)
rbp_cands_gr[, c("gene_id", "gene_name", "tag")]
```

Additionally, the transcript names and exon locations for
these genes are available.  With this dataset, we have
ENSG, ENST and ENSP identifiers.
```{r lktxex}
data(rbp_cands_txex)
rbp_cands_txex
```

# Differential expression patterns in RBP knockdown experiments

DESeq2 analyses are provided.  An example table is:
```{r lkdes2}
data(demoDE)
head(demoDE)
```

The `enhanceDE` utility will bind location and nomenclature
information.  The tables supplied by ENCORE are identified
only through their filenames, so an additional argument
must be supplied to give context by naming, informally, the
manipulated gene.  Here we verify
that the 'knockdown' of ZNF317 has the intended effect.
```{r lken}
dat = enhanceDE(demoDE, "ZNF317")
dat |> as.data.frame() |> dplyr::filter(gene_name == "ZNF317")
```
